---
title: "Assignment 3 Task 3"
author: "Kirsten Hodgson"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(kableExtra)

urchins <- read.csv(here("data","urchins.csv"))

urchins_clean <- urchins %>% clean_names()
```

## Size comparisons of purple and red urchins

#### Exploratory data visualization

1. Faceted histogram:

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = urchins_clean, aes(x = size)) + 
  geom_histogram(bins = 15) + #Create histogram of urchin size with 15 bins
  facet_wrap(~common_name) + #Facet so can clearly see histogram for both common names
  labs(x = "Urchin diameter (cm)", y = "Urchin count") #Add labels to axes
```

2. Box plot

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = urchins_clean, aes(x = common_name, y = size)) +
  geom_boxplot() + #Creates boxplot with common name on x-axis and size on y-axis
  labs(x = "Urchin Species", y = "Urchin diameter (cm)") #Labels axes
```

#### Descriptive statistics

```{r, echo = TRUE, message = FALSE, warning = FALSE}
urchins_summary <- urchins_clean %>% 
  group_by(common_name) %>% #Tell R to group by common name and then...
  summarize(mean_size = mean(size), #Find the mean for the two different common names...
            median_size = median(size), #The median...
            standard_dev = sd(size), #The standard deviation...
            sample_size = n()) 

kbl(urchins_summary,digits = 2, col.names = c("Common Name", #Round decimal digits to 2 and...
                                              "Mean Diameter (cm)", #Rename columns
                                              "Median Diameter (cm)", 
                                              "Standard Deviation", 
                                              "Sample Size")) %>%  
  kable_styling()
  
```

From the exploratory graphs and the summary statistics, the mean diameter of red urchins is larger than the mean diameter of purple urchins. The actual difference in mean diameter is 2.307 cm, which is a 41% difference compared to the average of the two means (which is 5.623). The red urchin population also has a larger median diameter (actual difference = 2.5 cm), and a higher standard deviation - so higher spread. The histograms both appear more or less normally distributed, though the purple urchin distribution does appear more positively skewed than the red urchin distribution.

## Purple urchins by site and year

#### Data wrangling

```{r, echo = TRUE, message = FALSE, warning = FALSE}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% #Keep only observation for purple urchins
  mutate(date_new = mdy(date), .keep = "unused") %>% #Make a new date column with class = Date
  mutate(year_new = year(date_new)) %>% #Add the year column
  mutate(site_full = case_when( 
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef")) #Add a column with entries depending on site
```

#### Summary table and exploratory graph

1. Summary table

```{r, echo = TRUE, message = FALSE, warning = FALSE}
urchins_site_year <- purple_urchins %>% 
  count(site_full, year_new) #Count purple urchin observations by site and year
```

2. Exploratory graph

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = urchins_site_year, aes(x = year_new, y =n)) + #Create graph of urchin count by year
  geom_col() + #As a column chart
  facet_wrap(~site_full) + #Faceted by site
  labs(x = "Year", y = "Count of purple urchins")
```

